Supplemental Information: Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite

### 2022-02-14 supp_v9.Rmd
some ms and prepping for final validation

---
title: "Supplemental Information: Using hydraulic geometry and the SWOT satellite to remotely sense river gas exchange velocity"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Peter A Raymond:
      institute: Yale
  - Merritt H Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
  - Yale: School of the Environment, Yale University, New Haven, CT
bibliography: References.bib
csl: agu_citation_style.csl
output:
  word_document:
      pandoc_args:
        - '--lua-filter=scholarly-metadata.lua'
        - '--lua-filter=author-info-blocks.lua'
      reference_docx: word-styles-reference.docx

@@

@@

## Table S3
*Table S3: Details on the 3 depth hydraulic geometry models used to estimate FCO2 from the SWOT rivers (section 2.4).*

|**Name**|**Depth equation**|**Velocity equation**|**Description**|**Reference**|
|--------|------------------|---------------------|---------------|-------------|
|Brinkerhoff 2019|$D=0.265Q^{0.395}$|$\bar{U}=0.284Q^{0.191}$|530,945 measurements made across the United States at streamgauges|this study; @brinkerhoffReconcilingAtaStationAtManyStations2019|
|Raymond 2012|$D=0.409Q^{0.294}$|$\bar{U}=0.194Q^{0.285}$||1,026 measurements across the United States|@raymondScalingGasTransfer2012|
|Raymond 2013|$D=0.409Q^{0.294},  D=0.449Q^{0.37}$|$\bar{U}=0.194Q^{0.285}, \bar{U}=0.346Q^{0.120}$|Average of the Raymond 2012 equation and one using 9,811 measurements at US streamgauges|@raymondGlobalCarbonDioxide2013|

>>>>

---
title: "Supplemental Information: Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Peter A Raymond:
      institute: Yale
  - Merritt H Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
  - Yale: School of the Environment, Yale University, New Haven, CT
bibliography: References.bib
csl: agu_citation_style.csl
output:
  word_document:
      pandoc_args:
        - '--lua-filter=scholarly-metadata.lua'
        - '--lua-filter=author-info-blocks.lua'
      reference_docx: word-styles-reference.docx

@@

## Text S4: Literature $CO_2$ flux models
Here, we describe in more detail the $CO_2$ flux models used to compare against BIKER in section 3.2.2. The specific equations used are outlined in Table S3.

In brief, gas fluxes ($FCO_2$) are only obtainable at the global scale via predictive equations for $k_{CO_2}$, as outlined in the main text. Here, we obtain $FCO_2$ using the in situ $CO_2$ data described in section 3.1 and $k_{CO_2}$ estimates calculated using 1) equation 7, 2) in situ streamflow data, and 3) hydraulic scaling equations used in the literature to predict river velocity and depth from discharge. Specifically, these literature equations include the 'Raymond 2012', 'Raymond 2013', and 'Brinkerhoff 2019' models.

For example, the 'Raymond 2012' model uses the hydraulic scaling equations from @raymondScalingGasTransfer2012 to predict velocity and depth, which are in turn used to predict $k_{CO_2}$, which finally is used to predict $FCO_2$. Similarly, the 'Raymond 2013' model uses the equations outlined in @raymondGlobalCarbonDioxide2013 and the 'Brinkerhoff 2019' model uses new equations fit to the hydraulics dataset from @brinkerhoffReconcilingAtaStationAtManyStations2019. We chose to include the 'Brinkerhoff 2019' model as the training dataset is far larger than those used in either of the previous two models (Table S3: 104,624 versus 10,837 versus 1,026, respectively). Finally, we converted from $k_600$ to $k_{CO_2}$ via @raymondScalingGasTransfer2012.

@@

## Table S3
*Table S3: Details on the 3 literature models for $CO_2$ fluxes. See Text S4 for more details.*

|**Name**|**Depth equation**|**Velocity equation**|**Description**|**Reference**|
|--------|------------------|---------------------|---------------|-------------|
|Brinkerhoff 2019|$D=0.265Q^{0.395}$|$\bar{U}=0.284Q^{0.191}$|104,624 measurements made across the United States at streamgauges|brinkerhoffReconcilingAtaStationAtManyStations2019|
|Raymond 2012|$D=0.409Q^{0.294}$|$\bar{U}=0.194Q^{0.285}$||1,026 measurements across the United States|@raymondScalingGasTransfer2012|
|Raymond 2013|$D=0.409Q^{0.294},  D=0.449Q^{0.37}$|$\bar{U}=0.194Q^{0.285}, \bar{U}=0.346Q^{0.120}$|Average of the Raymond 2012 equation and one using 9,811 measurements at US streamgauges|@raymondGlobalCarbonDioxide2013|
